
---
title: "Assignment 1, EDDA 2017"
author: "Arash and Fabio, group 22"
date: "5 April 2017"
output:
  html_document: default
  pdf_document: default
highlight: tango
fontsize: 11pt
---

## EXERCISE 1
If you place the file assign1.RData in your R-directory and type
> load(file="assign1.RData")
you will have data vectors x1, x2,..., x5 in your R working directory. Make a histogram and a QQ-plot for each of them, and decide which ones could have been sampled from a normal distribution. Experiment by simulating some normal samples of similar sizes and looking at their QQ-plots, before you make up your mind!

```{r}
load(file="assign1.RData")

```
### 1.1 Figures
```{r,fig.margin = TRUE,fig.width=6,fig.height=3}
par(mfrow=c(1,2))
hist(x1); qqnorm(x1); shapiro.test(x1)
hist(x2); qqnorm(x2);shapiro.test(x2)
hist(x3); qqnorm(x3);shapiro.test(x3)
hist(x4); qqnorm(x4);shapiro.test(x4)
hist(x5); qqnorm(x5);shapiro.test(x5)
```
comparing above graph to graphs of normal samples
the QQ plot shows linear corleation in a sample of 1000 from a normal distribution 
x3 and x4 are the 
From the five histogram and QQ-plots here above, we should be able to identify
whether # the data follows a normal distribution. a plot using qqnorm of a
sample from a normal distribution will #show approximately a straight line, and
a deviation from a line indicates that #the sample was not taken from a normal
population. We have that  x1, #x2, x3, x4 and x5 have sizes of 20, 100, 100, 30
and 40, respectively.

The elements x2, x3 and x4 look very much like a QQ-plot of #a normal
distribution. Even though x1 has an uprising behavior, it could still #have been
sampled from a non-normal distribution as it does not exactly follow #a straight
line behaviour. Finally, the x5 curve looks more like an exponential one rather
than #a straight line, thus it does not invite the thought that it has been
sampled #from a normal distribution. ### 1.2 Figures

```{r,fig.margin = TRUE,fig.width=6,fig.height=3}
x=rnorm(1000)
par(mfrow=c(1,2))
hist(x) ; qqnorm(x)
```
***

## Exercise 2

### 2.1 T-test

Below we perform a one sample t-test for the artificial date 
(that we generated ourselves).

```{r}
mu=0.2
x=rnorm(100,mu,1) # creating artificial data
t.test(x,mean=0)   # t.test(x,alternative=c("two.sided"),conf.level=0.95,mu=10)
```

 
Of course, we often do not need to report the whole output of R-commands, only certain
values of the output. For example, below we perform a two-smaple t-test and report 
only the value of t-statistics and the p-pavue.

```{r}
mu=0;nu=0.5
x=rnorm(50,mu,1); y=rnorm(50,nu,1) # creating artificial data
ttest=t.test(x,y) 
```

The value of t-statistics in the above evaluation is  `r round(ttest[[1]],2)` and the p-value is `r round(ttest[[3]],4)`.
  




#EXERCISE 2

```{r}
twoSampleT <- function(mu,nu,m,n,sd,pValue){
  B=1000
  p=numeric(B)
  c = 0  
  c10 =0; 
  for (b in 1:B) {
    x=rnorm(m,mu,sd)
    y=rnorm(n,nu,sd)
    p[b]=t.test(x,y,var.equal=TRUE)[[3]]
    if (p[b] < pValue){ c= c+1}
  }
  hist(p,main=c("with",pValue ))
}
```

```{r}
twoSampleTNU <- function(mu,nuMin,nuMax,nuStep,m,n,sd,pValue){
  B=1000
  p=numeric(B)
  power = numeric(B)
  c = 0  
  c10 =0
  for (nu in seq(nuMin,nuMax,by=nuStep)){
  for (b in 1:B) {
    x=rnorm(m,mu,sd)
    y=rnorm(n,nu,sd)
    p[b]=t.test(x,y,var.equal=TRUE)[[3]]
    if (p[b] < pValue){ c= c+1}
  }
  power[nu]=mean(p<0.05)
}
  hist(power,main=c("with",pValue ))
}
```

### 1. Set mu=nu=180, m=n=30 and sd=10. Repeat the script 1000 times, and record the 1000 p-values. How many p-values are smaller than 5%? How many are smaller than 10%? What is the distribution of the p-values (make a histogram)?

```{r}
par(mfrow=c(1,2))
print(twoSampleT(180,180,30,30,10,0.05))
print(twoSampleT(180,180,30,30,10,0.10))

```
### 2. Set mu=nu=180, m=n=30 and sd=1. Answer the same questions.
```{r}
par(mfrow=c(1,2))
print(twoSampleT(180,180,30,30,1,0.05))
print(twoSampleT(180,180,30,30,1,0.10))
```

###3. Set mu=180, nu=175, m=n=30 and sd=6. Answer the same questions.
```{r}
par(mfrow=c(1,2))
print(twoSampleT(180,175,30,30,6,0.05))
print(twoSampleT(180,175,30,30,6,0.10))
```
###4. Explain the findings.


#EXERCISE 3
###1. Set mu=180, m=n=30 and sd=5. Calculate the power of the t-test for every value of nu in the grid seq(175,185,by=0.1). Plot the power as a function of nu.
###2. Set mu=180, m=n=100 and sd=5. Repeat the preceding exercise. Add the plot to the preceding plot.
###3. Set mu=180, m=n=30 and sd=100. Repeat the preceding exercise.
###4. Explain the findings.
